---
title: "Delaware"
output: 
  rmdformats::downcute:
    toc_depth: 3
---


```{r message=FALSE, warning=FALSE, include=FALSE}
#Load packages we'll need throughout the entire document
library(readxl)
library(data.table)
library(DT)
library(sf)
library(mapview)
library(dplyr)
library(scales)
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(nngeo)
library(leaflet)
library(janitor)
library(DT)
library(kableExtra)
library(ggmap)
source('helper_functions.R')
library(RColorBrewer)

library(tidyr)
options(scipen = 100)

```



```{r message=FALSE, warning=FALSE, include=FALSE}

state_list <- c("IL", "MA", "MN", "NJ", "NM", "NY",  "CA")

hifld_df <- st_read("utility_zones/HIFLD/Electric_Retail_Service_Territories (1)/Electric_Retail_Service_Territories.shp") 

temp_util <- hifld_df %>%
  filter(ID %in% c(15473, 10817, 5701)) %>%
  mutate(new_name = case_when(
    ID == 15473 ~ "Public Service Company",
    ID == 10817 ~ "Xcel",
    ID == 5701 ~ "El Paso Electric"
  ))
census_api_key('ba3d95cae3658f3145daf7f66119ed58a52aaca3')

get_tracts <- get_acs(
  geography="tract", 
  state=state_list,
  variables=c("Median Income" = "S1901_C01_012E"),
  year=2017, 
  geometry=TRUE) %>%
  mutate(MedInc = estimate) %>%
  select(geoid=GEOID, geometry)


```



# test

some other stuff 

# test

some stuff 

## test tes
(summarized 8/30/2023)



* Gross annual income, by family size, is at or below 200% of the Federal Poverty Guidelines, or 60% of the state median household income published by the United States Census Bureau, whichever is greater.


(from SB 2)


* Low-income Customer verification (*includes self-attestation*):
  * The CEF shall require income data verification to determine eligibility for Low-income Customers. Proof of eligibility required for Low-income Customers shall include:
    * (1) Proof of income of the account holder (e.g., pay stub or W2)
    * (2) proof of participation in a low income discount program including Medicaid, SSI, TANF, GA, WIC, LIHEAP, SNAP or food stamps
    * (3) proof that the Low-income Customer lives in a census block where the median household income is at or below 200% of the U.S. Federal Poverty Guidelines or 60% of the state median household income published by the United States Census Bureau, whichever is greater
    * (4) by living in a low-income master-metered building
    * (5) A written attestation by the Low-income Customer that their total household income is at or below 200% of the Federal Poverty Guidelines, or 60% of the state median household income published by the United States Census bureau, whichever is greater
    * NOTE For owners or operators of master-metered buildings, a written attestation that their tenants meet the income eligibility requirements for Low-income Customers

