---
title: "Massachusetts"
output: 
  rmdformats::downcute:
    toc_depth: 3
---


---
```{r setup, include=FALSE}

source('helper_functions.R')
load_packages()



```


# Overivew





```{r message=FALSE, warning=FALSE, include=FALSE}

get_tracts <- get_acs(
  geography="tract", 
  state="MA",
  variables=c("Median Income" = "S1901_C01_012E"),
  year=2019, 
  geometry=TRUE) %>%
  mutate(MedInc = estimate) %>%
  select(geoid=GEOID, geometry)

temp_util <- st_read("utility_zones/MA/pubutil/TOWNS_POLY_V_ELEC.shp") %>%
  mutate(
    new_name = case_when(
      grepl("Municipal", ELEC_LABEL) ~ "Municipal",
      grepl("NSTAR", ELEC_LABEL) ~ "NSTAR",
      TRUE ~ ELEC_LABEL
    )
  )

ma_ej_blocks <- st_read("data/MA/EJ_POLY.shp")


```






```{r message=FALSE, warning=FALSE, include=FALSE}
chas_df <- load_chas() %>%
    mutate(State_Name = case_when(
    state == 17 ~ "Illinois",
    state == 25 ~ "Massachusetts",  
    state == 27 ~ "Minnesota",  
    state == 34 ~ "New Jersey",
    state == 35 ~ "New Mexico",  
    state == 36 ~ "New York",
    state == 06 ~ "California",
    TRUE ~ "Other"
  )) %>%
  filter(State_Name == "Massachusetts")



add_geo <- merge(chas_df, get_tracts) 

add_geo <- st_as_sf(add_geo)


temp_util <- st_transform(temp_util, st_crs(add_geo)) %>%
  st_make_valid()

temp_util <- temp_util %>%
  group_by(new_name) %>%
  summarize(geometry = st_union(geometry), .groups = "drop")

```




```{r echo=FALSE, message=FALSE, warning=FALSE}

pal5 <- colorFactor(
  palette="viridis",
  domain=temp_util$new_name
)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=temp_util,
              group="Utility Zones",
              stroke=TRUE,
              color=~pal5(temp_util$new_name),
              dashArray="3",
              fillOpacity=0.5,
              popup=paste("Utility Zone: ", temp_util$new_name)) %>%
  addLegend(position="bottomright",
            pal=pal5,
            values=temp_util$new_name,
            title="Solstice MA Utilities")
```

# Geoqualification

The below map shows the qualified Environmental Justice (EJ) 2020 census block groups for Massachusetts. Note, the below represent block groups that are tagged with the presence of an EJ metric, and does not delineate between EJ classifications (Income, Minority, Language, combination of the three). 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# temp <- ma_ej_blocks %>%
#   st_transform(st_crs(get_tracts))
# 
# 
# st_write(temp, "/Users/jacobford/Library/CloudStorage/GoogleDrive-jake@solstice.us/Shared drives/Product | Jake/Geocoding/MA/MA_2020_EJ_Blocks.geojson", append=FALSE)

mapview(ma_ej_blocks)
```

# LMI Qualificaiton






```{r message=FALSE, warning=FALSE, include=FALSE}
ma_tracts <- get_acs(
  geography="tract", 
  state="MA",
  variables=c("Total_Pop" = "B01001_001","snap_hh" = "B22001_002","snap_pop" = "B19058_002"
              
              ),
  year=2019, 
  geometry=TRUE) %>%
  group_by(GEOID) %>%
  summarize(total_pop = estimate[variable=="Total_Pop"],
            snap_hh = estimate[variable=="snap_hh"],
            snap_pop = estimate[variable=="snap_pop"])







total_ma_tracts <- cbind(add_geo, ma_tracts)

```


# LIFT Solar

There are 145 community solar projects in New Jersey according to the LIFT solar database. 



```{r echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE}

lift_df <- read.csv("data/groundswell_lift/cs-projects-2023-08-18.csv") 

lift_df <- separate(lift_df, GeoCode, into = c("long", "lat"), sep = ",")

lift_df <- st_as_sf(lift_df, coords = c("long", "lat"), crs = st_crs(get_tracts))



temp_df <- add_geo %>%
  filter(State_Name == "Massachusetts") 

temp_lift <- lift_df %>%
  filter(State == "Massachusetts")


pal_potLMI <- colorNumeric(
  palette = "viridis",
 # reverse=TRUE,
  domain = temp_lift$Potential...LMI.Subscribers)

pal_cap <- colorNumeric(
  palette = "viridis",
 # reverse=TRUE,
  domain = temp_lift$Project.Capacity.KW.AC)



pal <- colorNumeric(
  palette = "viridis",
 # reverse=TRUE,
  domain = temp_df$AMI_80)

pal2 <- colorNumeric(
  palette = "Purples",
 # reverse=TRUE,
  domain = temp_df$AMI_80_Pct)

pal3 <- colorNumeric(
  palette = "viridis",
 # reverse=TRUE,
  domain = total_ma_tracts$snap_hh)

library(tidyr)





map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=temp_util,
              group="Utility Zones",
              stroke=TRUE,
              color="black",
              dashArray="3",
              fillOpacity=0,
              popup=paste("Utility Zone: ", temp_util$comp_short)
  ) %>%
  
  addPolygons(
    data = temp_df,
    group="80% AMI",  # This should be consistent with the group name in the addLegend function
    stroke=FALSE,
    smoothFactor=0.2,
    fillOpacity=0.85,
    color=~pal(AMI_80),
    popup=paste("Tract: ", temp_df$geoid, "<br>", 
                "HHs at 80% AMI: ", temp_df$AMI_80)
  ) %>%

  addLegend("bottomleft",
            group="80% AMI",  # This should be consistent with the group name in the addPolygons function
            pal = pal,
            values = temp_df$AMI_80,
            title="Number of HHs at 80% AMI"
  ) %>%
  
  # addPolygons(
  #   data=nm_df,
  #   group="80% AMI Percent",  # This should be consistent with the group name in the addLegend function
  #   stroke=FALSE,
  #   smoothFactor=0.2,
  #   fillOpacity=0.7,
  #   color=~pal2(AMI_80_Pct),
  #   popup=paste("Tract: ", nm_df$geoid, "<br>", 
  #               "Percent of HHs at 80% AMI: ", nm_df$AMI_80_Pct)
  # ) %>%
  # 
  # addLegend("bottomleft",
  #           group="80% AMI Percent",  # This should be consistent with the group name in the addPolygons function
  #           pal = pal2,
  #           values = nm_df$AMI_80_Pct,
  #           title="Percent of HHs at 80% AMI"
  # ) %>%
  
  addPolygons(
    data=total_ma_tracts,
    group="SNAP",
    stroke=FALSE,
    smoothFactor=0.2,
    fillOpacity=0.7,
    color=~pal3(snap_hh),
    popup=paste("Tract: ", total_ma_tracts$GEOID, "<br>",
                "Number of HHs Receiving SNAP: ", total_ma_tracts$snap_hh)
  ) %>%

  addLegend("bottomleft",
            group="SNAP",
            pal = pal3,
            values = total_ma_tracts$snap_hh,
            title="Number of HHs Receiving SNAP:"
  ) %>%
  
    addCircleMarkers(data = temp_lift,
                   group="Potential LMI Subscribers",
                   popup = ~paste("Program Name: ", Program.Name, "<br>",
                                  "Developer/Owner: ", Project.Developer.or.Owner, "<br>", 
                                  "Project Capacity: ", Project.Capacity.KW.AC, "<br>",
                                  "LMI Savings: ", LMI.Customer.Savings.., "<br>",
                                  "Potential LMI Subscribers: ", Potential...LMI.Subscribers),
                   radius = sqrt(temp_lift$Potential...LMI.Subscribers), 
                   color = ~pal_potLMI(Potential...LMI.Subscribers)) %>%
  addLegend("bottomright",
            group="Potential LMI Subscribers",  # This should be consistent with the group name in the addPolygons function
            pal = pal_potLMI,
            values = temp_lift$Potential...LMI.Subscribers,
            title="Potential LMI Subscribers") %>%

    ## Project Capacity ## 
  
  addCircleMarkers(data = temp_lift,
                   group="Project Capacity",
                   popup = ~paste("Program Name: ", Program.Name, "<br>",
                                  "Developer/Owner: ", Project.Developer.or.Owner, "<br>", 
                                  "Project Capacity: ", Project.Capacity.KW.AC, "<br>",
                                  "LMI Savings: ", LMI.Customer.Savings.., "<br>",
                                  "Potential LMI Subscribers: ", Potential...LMI.Subscribers),
                   radius = (temp_lift$Project.Capacity.KW.AC), 
                   color = ~pal_cap(Project.Capacity.KW.AC)) %>%
  addLegend("topleft",
             group="Project Capacity",  # This should be consistent with the group name in the addPolygons function
            pal = pal_cap,
            values = temp_lift$Project.Capacity.KW.AC,
            title="Project Capacity") %>%


  addLayersControl(
    overlayGroups=c( "80% AMI", "Utility Zones", "SNAP","Potential LMI Subscribers", "Project Capacity"),  # Update the order for consistency
    options = layersControlOptions(collapsed = FALSE)
  )

map %>%
 hideGroup("SNAP") %>% hideGroup("80% AMI") %>% hideGroup("Project Capacity")

```

