---
title: "New Mexico"
output: 
  rmdformats::downcute:
    toc_depth: 3
---


```{r setup, include=FALSE}

source('helper_functions.R')
load_packages()


```




```{r message=FALSE, warning=FALSE, include=FALSE}

state_list <- c("IL", "MA", "MN", "NJ", "NM", "NY",  "CA")

hifld_df <- st_read("utility_zones/HIFLD/Electric_Retail_Service_Territories (1)/Electric_Retail_Service_Territories.shp") 

temp_util <- hifld_df %>%
  filter(ID %in% c(15473, 10817, 5701)) %>%
  mutate(new_name = case_when(
    ID == 15473 ~ "Public Service Company",
    ID == 10817 ~ "Xcel",
    ID == 5701 ~ "El Paso Electric"
  ))
census_api_key('ba3d95cae3658f3145daf7f66119ed58a52aaca3')

get_tracts <- get_acs(
  geography="tract", 
  state=state_list,
  variables=c("Median Income" = "S1901_C01_012E"),
  year=2017, 
  geometry=TRUE) %>%
  mutate(MedInc = estimate) %>%
  select(geoid=GEOID, geometry)


```




# Geoqualification

In New Mexico, there is no geographic qualification methods, so we only focus on estimating the potential LMI population by various methods used in the state program. 




```{r message=FALSE, warning=FALSE, include=FALSE}
chas_df <- load_chas() %>%
    mutate(State_Name = case_when(
    state == 17 ~ "Illinois",
    state == 25 ~ "Massachusetts",  
    state == 27 ~ "Minnesota",  
    state == 34 ~ "New Jersey",
    state == 35 ~ "New Mexico",  
    state == 36 ~ "New York",
    state == 06 ~ "California",
    TRUE ~ "Other"
  ))



add_geo <- merge(chas_df, get_tracts) 

add_geo <- st_as_sf(add_geo)

nm_geo <- add_geo %>%
  filter(State_Name=="New Mexico")

temp_util <- st_transform(temp_util, st_crs(nm_geo)) %>%
  st_make_valid()
```

# LMI Qualifications


New Mexico LMI Qualification methods include:


  * Medicaid
  * SNAP
  * LIHEAP
  * First-time home owner programs
  * Affordable housing facilities
    - HUD data ([potential example](https://resources.hud.gov/#layers-menu))
    - State specific ([potential example](https://housingnm.org/find-housing/rentals/affordable))
  * State and federal income tax credit
  * Entire multi-family housing project may qualify if the entire load can be proved to be low-income subscribers, with consent of all tenants of record.
  * **LMI - 80% of AMI**
  

Census data is available for the population receiving Medicaid, SNAP, and number of households at or below 80% AMI. The following analysis quantifies the total eligible LMI population by each of these three categories.

Solstice operates in the following three utility zones in New Mexico:

  * Public Service Company of New Mexico
  * Excel
  * El Paso Electric

```{r echo=FALSE, message=FALSE, warning=FALSE}

pal5 <- colorFactor(
  palette="viridis",
  domain=temp_util$new_name
)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=temp_util,
              group="Utility Zones",
              stroke=TRUE,
              color=~pal5(temp_util$new_name),
              dashArray="3",
              fillOpacity=0.5,
              popup=paste("Utility Zone: ", temp_util$new_name)) %>%
  addLegend(position="bottomright",
            pal=pal5,
            values=temp_util$new_name,
            title="Solstice NM Utilities")
```
  


```{r message=FALSE, warning=FALSE, include=FALSE}

  # * Medicaid
  #   - total count of population with Medicaid- C27007_004 + C27007_007 + C27007_010 + C27007_014 + C27007_017 + C27007_020
  #   - geo: tract
  #   
  # * SNAP
  #   - total count of HH received Food Stamps/SNAP in last 12 months - B22001_002
  #   - total count of population received Food Stamps/SNAP in last 12 months - B19058_002
  #   - geo: tract
  # * LIHEAP
  #   - use this to calculate income by self attestation? 
  #   https://www.acf.hhs.gov/sites/default/files/documents/ocs/comm_liheap_im2002smiattachment_fy2021.pdf
  #   
  # * first-time home owner programs
  # * affordable housing facilities
  #   - HUD data: https://resources.hud.gov/#layers-menu
  #   - state specific to supplement: https://housingnm.org/find-housing/rentals/affordable
  # * low-income housing
  # * state and federal income tax credit
  # * entire multi-family housing project may qualify if the entire load can be proved to be low-income subscribers, with consent of all tenants of record.
  # 
  # LMI - 80% of AMI from HUD..
# acs_vars <- load_variables(2019, "acs5")

nm_blocks <- get_acs(
  geography="tract", 
  state="NM",
  variables=c("snap_hh" = "B22001_002",
              "snap_pop" = "B19058_002",
              "medicaid_1" = "C27007_004",
              "medicaid_2" = "C27007_007",
              "medicaid_3" = "C27007_010",
              "medicaid_4" = "C27007_014",
              "medicaid_5" = "C27007_017",
              "medicaid_6" = "C27007_020",
              "Total_Pop" = "B01001_001"
              
              ),
  year=2019, 
  geometry=TRUE) %>%
  group_by(GEOID) %>%
  summarize(snap_hh = estimate[variable=="snap_hh"],
            medicaid_pop =estimate[variable=="medicaid_1"] + 
                          estimate[variable=="medicaid_2"] +
                          estimate[variable=="medicaid_3"] +
                          estimate[variable=="medicaid_4"] +
                          estimate[variable=="medicaid_5"] +
                          estimate[variable=="medicaid_6"] ,
            total_pop = estimate[variable=="Total_Pop"])

total_nm_tracts <- cbind(nm_geo, nm_blocks)


```


The table below shows total population by LMI qualification method by utility area coverage. Note, these LMI groups are not mutually exclusive, rather are snap shots of the individual method's population. 

```{r echo=FALSE, message=FALSE, warning=FALSE}


temp <- total_nm_tracts %>% 
  st_point_on_surface() %>%
  st_make_valid() %>%
  st_join(temp_util %>% select(NAME)) %>%
  st_drop_geometry() %>%
  group_by(Name = NAME) %>%
  filter(!is.na(Name)) %>%
  summarize(`Pop under 80 AMI` = sum(AMI_80),
            `Medicaid` = sum(medicaid_pop),
            `SNAP HH` = sum(snap_hh),
            `Total Population` = sum(total_pop)) %>%
  arrange(desc(`Pop under 80 AMI`)) 

datatable(temp, caption = "New Mexico, Sources: ACS 2019 and HUD CHAS")%>%
  formatCurrency('Pop under 80 AMI',currency = "", interval = 3, mark = ",") %>%
  formatCurrency('Medicaid',currency = "", interval = 3, mark = ",") %>%
  formatCurrency('SNAP HH',currency = "", interval = 3, mark = ",") %>%
  formatCurrency('Total Population',currency = "", interval = 3, mark = ",") 





```


The map shows the census tracts loaded with the LMI qualification methods by utility zone coverage. 




```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=8}

nm_df <- add_geo %>%
  filter(State_Name == "New Mexico") 

pal <- colorNumeric(
  palette = "Reds",
 # reverse=TRUE,
  domain = nm_df$AMI_80)

pal2 <- colorNumeric(
  palette = "Purples",
 # reverse=TRUE,
  domain = nm_df$AMI_80_Pct)

pal3 <- colorNumeric(
  palette = "Greens",
 # reverse=TRUE,
  domain = nm_blocks$snap_hh)

pal4 <- colorNumeric(
  palette = "Blues",
 # reverse=TRUE,
  domain = nm_blocks$medicaid_pop)


 



map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data=temp_util,
              group="Utility Zones",
              stroke=TRUE,
              color="black",
              dashArray="3",
              fillOpacity=0,
              popup=paste("Utility Zone: ", temp_util$new_name)
  ) %>%
  
  addPolygons(
    data = nm_df,
    group="80% AMI",  # This should be consistent with the group name in the addLegend function
    stroke=FALSE,
    smoothFactor=0.2,
    fillOpacity=0.7,
    color=~pal(AMI_80),
    popup=paste("Tract: ", nm_df$geoid, "<br>", 
                "HHs at 80% AMI: ", nm_df$AMI_80)
  ) %>%

  addLegend("bottomleft",
            group="80% AMI",  # This should be consistent with the group name in the addPolygons function
            pal = pal,
            values = nm_df$AMI_80,
            title="Number of HHs at 80% AMI"
  ) %>%
  
  # addPolygons(
  #   data=nm_df,
  #   group="80% AMI Percent",  # This should be consistent with the group name in the addLegend function
  #   stroke=FALSE,
  #   smoothFactor=0.2,
  #   fillOpacity=0.7,
  #   color=~pal2(AMI_80_Pct),
  #   popup=paste("Tract: ", nm_df$geoid, "<br>", 
  #               "Percent of HHs at 80% AMI: ", nm_df$AMI_80_Pct)
  # ) %>%
  # 
  # addLegend("bottomleft",
  #           group="80% AMI Percent",  # This should be consistent with the group name in the addPolygons function
  #           pal = pal2,
  #           values = nm_df$AMI_80_Pct,
  #           title="Percent of HHs at 80% AMI"
  # ) %>%
  
  addPolygons(
    data=nm_blocks,
    group="SNAP",
    stroke=FALSE,
    smoothFactor=0.2,
    fillOpacity=0.7,
    color=~pal3(snap_hh),
    popup=paste("Tract: ", nm_blocks$GEOID, "<br>", 
                "Number of HHs Receiving SNAP: ", nm_blocks$snap_hh)
  ) %>%
  
  addLegend("bottomleft",
            group="SNAP",
            pal = pal3,
            values = nm_blocks$snap_hh,
            title="Number of HHs Receiving SNAP:"
  ) %>%
  
  addPolygons(
    data=nm_blocks,
    group="Medicaid",
    stroke=FALSE,
    smoothFactor=0.2,
    fillOpacity=0.7,
    color=~pal4(medicaid_pop),
    popup=paste("Tract: ", nm_blocks$GEOID, "<br>", 
                "Number of HHs Receiving Medicaid: ", nm_blocks$medicaid_pop)
  ) %>%
  
  addLegend("bottomleft",
            group="Medicaid",
            pal = pal4,
            values = nm_blocks$medicaid_pop,
            title="Number of HHs Receiving Medicaid:"
  ) %>%
   
  addLayersControl(
    overlayGroups=c( "80% AMI", "Utility Zones", "SNAP", "Medicaid"),  # Update the order for consistency
    options = layersControlOptions(collapsed = FALSE)
  )

map %>%
 hideGroup("SNAP") %>% hideGroup("80% AMI") %>% hideGroup("Medicaid")

```







## Low Income Housing

To estimate the total number of New Mexicans residing in low income housing, we first require a list of eligible properties. We take each affordable rental property listed on the [MFA Housing New Mexico](https://housingnm.org/find-housing/rentals/affordable) website. After geocoding each point, we have several options for determining the population concentration around each point. A common technique in transportation modeling is applying an isocrhone, or a travel time map. In our case, we want to see how many eligible population of SNAP, Medicaid and 80% AMI households are within a 15 and 30 minute driving radius around each point. 

```{r message=FALSE, warning=FALSE, include=FALSE}
# nm_data <- read_excel("data/nm/nm_public_housing.xlsx")
# 
# 
# 
# library(ggmap)
# 
# # Assuming your 'nm_data' dataframe has columns 'Address' and 'Address 2'
# # Create a new column to store the full address (combining Address and Address 2)
# nm_data$full_address <- paste(nm_data$Address, nm_data$`Address 2`, sep = ", ")
# 
# # Use geocode() to get latitude and longitude for each address
# geocoded_data <- geocode(nm_data$full_address)
# 
# # Add latitude and longitude columns to the original dataframe
# nm_data$Latitude <- geocoded_data$lat
# nm_data$Longitude <- geocoded_data$lon
# 
# nm_data <- nm_data %>%?
#   filter(!is.na(Longitude))
# 
# nm_data <- st_as_sf(nm_data, coords = c("Longitude", "Latitude"), crs = st_crs(nm_blocks))

# st_write(nm_data, "nm_public_housing.shp")




nm_public_housing <- st_read("data/nm/nm_public_housing.shp")

# nm_public_housing_isos <- nm_public_housing %>%
#   mb_isochrone(time = 30, profile = "driving", id_column = "Name",
#                access_token = "blahblahblah"

nm_public_housing_isos_15 <- st_read("isochrones/nm_public_housing_isos_15.shp") 
nm_public_housing_isos_15 <- st_transform(nm_public_housing_isos_15, st_crs(nm_public_housing))

nm_public_housing_isos_30 <- st_read("isochrones/nm_public_housing_isos_30.shp")
nm_public_housing_isos_30 <- st_transform(nm_public_housing_isos_30, st_crs(nm_public_housing))


```

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(leaflet.extras)


nm_pub_in_util <- nm_public_housing %>%
  st_point_on_surface() %>%
  st_join(temp_util %>% select(util_name = NAME)) %>%
  filter(!is.na(util_name))


houses_in_utilies <- nm_pub_in_util$Name

iso_15 <- nm_public_housing_isos_15 %>%
  filter(id %in% houses_in_utilies)

iso_30 <- nm_public_housing_isos_30 %>%
  filter(id %in% houses_in_utilies)


map %>%
  # addMarkers(data=nm_public_housing,
  #                  group="Public Housing",
  #               #   color="black",
  #               #   radius=2,
  #                  popup=paste("Name: ", nm_public_housing$Name)) %>%
  addMarkers(data = nm_pub_in_util,
             group="Public Housing",
           #  color="red",
            popup=paste("Name: ", nm_pub_in_util$util_name)) %>%
  addPolygons(data =iso_15,
              group = "15 Minute Driving",
              fillOpacity = 0.2,
              color = "orange") %>%
    addPolygons(data =iso_30,
              group = "30 Minute Driving",
              fillOpacity = 0.2,
              color = "darkgreen") %>%

  addLayersControl(
    overlayGroups=c("80% AMI", "Utility Zones", "SNAP", "Medicaid", "Public Housing", "15 Minute Driving", "30 Minute Driving"),  # Update the order for consistency
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup("80% AMI Percent") %>% hideGroup("SNAP") %>% hideGroup("80% AMI") %>% hideGroup("Medicaid") 


```





```{r message=FALSE, warning=FALSE, include=FALSE}



# Assuming 'snap' column is present in total_nm_tracts
# Check validity of isochrones
iso_15_valid <- st_is_valid(iso_15)


# Find indices of invalid isochrones
invalid_indices <- which(!iso_15_valid)


# Filter out invalid isochrones
valid_iso_15 <- iso_15[-invalid_indices, ]

total_nm_tracts$total_area_tract <- st_area(total_nm_tracts)


# Find the intersection of valid_iso_15 with total_nm_tracts
intersecting_tracts <- st_intersection(total_nm_tracts, valid_iso_15)

# Calculate the area of valid_iso_15 that covers each tract
intersecting_tracts$covered_area <- st_area(intersecting_tracts)

# Calculate the proportion of coverage for each tract
intersecting_tracts$proportion_coverage <- intersecting_tracts$covered_area / total_nm_tracts$total_area_tract

# If you want to consider the 'snap' column for calculating the proportion
# You can adjust the calculation based on your needs
intersecting_tracts$proportion_snap_coverage <- intersecting_tracts$proportion_coverage * intersecting_tracts$snap_hh
intersecting_tracts$proportion_medicaid_pop <- intersecting_tracts$proportion_coverage * intersecting_tracts$medicaid_pop
intersecting_tracts$proportion_80ami_coverage <- intersecting_tracts$proportion_coverage * intersecting_tracts$AMI_80
intersecting_tracts$proportion_total_pop <- intersecting_tracts$proportion_coverage * intersecting_tracts$total_pop


temp_15 <- intersecting_tracts %>%
  st_drop_geometry() %>%
  group_by(geoid) %>%
  summarize(total_snap = max(snap_hh), 
            total_medicaid = max(medicaid_pop),
            total_80ami = max(AMI_80),
            covered_area = sum(proportion_coverage),
            total_pop = max(total_pop)) %>%
  mutate(covered_prop = 
            case_when(
                as.numeric(covered_area) >= 100 ~ 1,
                TRUE ~ as.numeric(covered_area))
    
    
    ) %>%
  mutate(
    proportion_snap = covered_prop * total_snap/100,
    proportion_medicaid = covered_prop * total_medicaid/100,
    proportion_80ami = covered_prop * total_80ami/100,
    proportion_total_pop = covered_prop * total_pop/100,
    isochrone = "15 Minute"
  ) %>%
  group_by(isochrone) %>%
  summarize(SNAP = sum(proportion_snap),
            
            Medicaid = sum(proportion_medicaid),
            LMI = sum(proportion_80ami),
            Total_Pop = sum(proportion_total_pop))



temp_15





```



```{r message=FALSE, warning=FALSE, include=FALSE}


# Assuming 'snap' column is present in total_nm_tracts
# Check validity of isochrones
iso_30_valid <- st_is_valid(iso_30)


# Find indices of invalid isochrones
invalid_indices <- which(!iso_30_valid)


# Filter out invalid isochrones
valid_iso_30 <- iso_30[-invalid_indices, ]

total_nm_tracts$total_area_tract <- st_area(total_nm_tracts)


# Find the intersection of valid_iso_30 with total_nm_tracts
intersecting_tracts <- st_intersection(total_nm_tracts, iso_30)

# Calculate the area of valid_iso_30 that covers each tract
intersecting_tracts$covered_area <- st_area(intersecting_tracts)

# Calculate the proportion of coverage for each tract
intersecting_tracts$proportion_coverage <- intersecting_tracts$covered_area / total_nm_tracts$total_area_tract

# If you want to consider the 'snap' column for calculating the proportion
# You can adjust the calculation based on your needs
intersecting_tracts$proportion_snap_coverage <- intersecting_tracts$proportion_coverage * intersecting_tracts$snap_hh
intersecting_tracts$proportion_medicaid_pop <- intersecting_tracts$proportion_coverage * intersecting_tracts$medicaid_pop
intersecting_tracts$proportion_80ami_coverage <- intersecting_tracts$proportion_coverage * intersecting_tracts$AMI_80
intersecting_tracts$proportion_total_pop <- intersecting_tracts$proportion_coverage * intersecting_tracts$total_pop


temp_30 <- intersecting_tracts %>%
  st_drop_geometry() %>%
  group_by(geoid) %>%
  summarize(total_snap = max(snap_hh), 
            total_medicaid = max(medicaid_pop),
            total_80ami = max(AMI_80),
            covered_area = sum(proportion_coverage),
            total_pop = max(total_pop)) %>%
  mutate(covered_prop = 
            case_when(
                as.numeric(covered_area) >= 100 ~ 1,
                TRUE ~ as.numeric(covered_area))
    
    
    ) %>%
  mutate(
    proportion_snap = covered_prop * total_snap/100,
    proportion_medicaid = covered_prop * total_medicaid/100,
    proportion_80ami = covered_prop * total_80ami/100,
    proportion_total_pop = covered_prop * total_pop/100,
    isochrone = "30 Minute"
  ) %>%
  group_by(isochrone) %>%
  summarize(SNAP = sum(proportion_snap),
            
            Medicaid = sum(proportion_medicaid),
            LMI = sum(proportion_80ami),
            Total_Pop = sum(proportion_total_pop))



temp_30

temp <- rbind(temp_15, temp_30)

```





```{r echo=FALSE, message=FALSE, warning=FALSE}

datatable(temp, caption = "New Mexico Distance from Public Housing and Selected Characteristics")%>%
  formatCurrency('LMI',currency = "", interval = 3, mark = ",") %>%
  formatCurrency('Medicaid',currency = "", interval = 3, mark = ",") %>%
  formatCurrency('SNAP',currency = "", interval = 3, mark = ",")  %>%
  formatCurrency('Total_Pop', currency="", interval = 3, mark = ",")



```




# LIFT Solar

There are zero community solar projects in New Mexico according to the LIFT solar database. 




