---
title: "Court Ordered Districts - More Fair?"
description: |
 North Carolina's New Congressional Districts and ML
author:
  - name: Jacob Ford
    url: https://jford-dchcmpo.github.io/Blog/posts/2021-12-30-Redistricting-and-Racial-Targetting/
date: 02-24-2022
output:
  distill::distill_article:
    self_contained: false
---


Deja-vu! The North Carolina Supreme Court rejected the maps that I previously analyzed, instead opting to use [maps](https://www.ncleg.gov/Redistricting) drawn by outside experts. Wowza! Let's, hopefully one last time, see how these maps compare to the 2010-2019 maps, the initially suggested redistricting by the General Assembly, the February update post the initial legal challenge, and now the court ordered drafted lines. 


```{r setup, include=FALSE}
##load packages and district lines

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#library(caliperR)
library(nngeo)
#library(tcadr)
library(sf)
library(leaflet)
library(PL94171)
library(janitor)
library(dplyr)
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(sf)
library(data.table)
library(dplyr)
library(plotly)
library(randomForest)

# remotes::install_github("walkerke/tidycensus")
# remotes::install_github("walkerke/tigris")

library(tigris)

vars_2020 <- load_variables(2020,"pl",cache=TRUE)
vars_2010 <- load_variables(2010,"sf1", cache=TRUE)
vars_2019 <- load_variables(2019,"acs5", cache=TRUE)

trm_taz_g2 <- st_read("/Users/jacobford/Library/Mobile Documents/com~apple~CloudDocs/R/Files/TRM_G2_TAZ_CaliperUpdated.shp")

nc_congressional_2019 <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts/2019 NC Congressional/NC_Congress_2019.shp") 
nc_congressional_2021 <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//2021 NC Congressional/SL 2021-174 Congress.shp")
nc_congressional_2021_feb <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//February Update/SL 2022-3.shp")
nc_congress_court <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//Final Court Ordered/Interim Congressional.shp")

house_old<- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//NC_House_2019.shp")
house_new<- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//NC_House_2021.shp")
house_feb <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//February Update/SL 2022-4.shp")

senate_old<- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//NC_Senate_2019.shp")
senate_new<- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//NC_Senate_2021.shp")
senate_feb <- st_read("/Users/jacobford/Documents/GitHub/Blog/Voting Districts//February Update/SL 2022-2.shp")

nc_congressional_2021$DISTRICT <- as.integer(nc_congressional_2021$DISTRICT)


##Change CRS


nc_congressional_2019 <- st_transform(nc_congressional_2019, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congressional_2021 <- st_transform(nc_congressional_2021, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congressional_2021_feb <- st_transform(nc_congressional_2021_feb, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congress_court <- st_transform(nc_congress_court, st_crs(trm_taz_g2)) %>% st_make_valid()


house_old <- st_transform(house_old, st_crs(trm_taz_g2)) %>% st_make_valid()
house_new <- st_transform(house_new, st_crs(trm_taz_g2)) %>% st_make_valid()
house_feb <- st_transform(house_new, st_crs(trm_taz_g2)) %>% st_make_valid()

senate_old <- st_transform(senate_old, st_crs(trm_taz_g2)) %>% st_make_valid()
senate_new <- st_transform(senate_new, st_crs(trm_taz_g2)) %>% st_make_valid()
senate_feb <- st_transform(senate_new, st_crs(trm_taz_g2)) %>% st_make_valid()




nc_congressional_2021 <- nc_congressional_2021[order(nc_congressional_2021$DISTRICT),]

nc_congressional_2021$Partisan_Lean <- c(-18,1,-20,-10,25,42,-20,-20,45,-26,-16,-16,-25,-12)

##source for paritisan lean: https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/
nc_congressional_2021$color <- case_when(
       nc_congressional_2021$Partisan_Lean  < -19 ~ -3,
        nc_congressional_2021$Partisan_Lean < -10 ~ -2,
        nc_congressional_2021$Partisan_Lean < -5 ~ -1,
        nc_congressional_2021$Partisan_Lean < 5 ~ 0, 
        nc_congressional_2021$Partisan_Lean  < 10 ~ 1,
        nc_congressional_2021$Partisan_Lean > 10 ~ 2 
         
)

nc_congressional_2021$Pred_Party <- case_when(
       nc_congressional_2021$Partisan_Lean  < 2 ~ 0,
        nc_congressional_2021$Partisan_Lean > 2 ~ 1
         
)

nc_congressional_2021_feb <- nc_congressional_2021_feb[order(nc_congressional_2021_feb$DISTRICT),]

nc_congressional_2021_feb$Partisan_Lean <- c(-18,1,-20,-10,25,42,-20,-20,45,-26,-16,-16,-25,-12)

##source for paritisan lean: https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/
nc_congressional_2021_feb$color <- case_when(
       nc_congressional_2021_feb$Partisan_Lean  < -19 ~ -3,
        nc_congressional_2021_feb$Partisan_Lean < -10 ~ -2,
        nc_congressional_2021_feb$Partisan_Lean < -5 ~ -1,
        nc_congressional_2021_feb$Partisan_Lean < 5 ~ 0, 
        nc_congressional_2021_feb$Partisan_Lean  < 10 ~ 1,
        nc_congressional_2021_feb$Partisan_Lean > 10 ~ 2 
         
)

nc_congressional_2021_feb$Pred_Party <- case_when(
       nc_congressional_2021_feb$Partisan_Lean  < 2 ~ 0,
        nc_congressional_2021_feb$Partisan_Lean > 2 ~ 1
         
)



nc_congress_court <- nc_congress_court[order(nc_congress_court$DISTRICT),]

nc_congress_court$Partisan_Lean <- c(-18,1,-20,-10,25,42,-20,-20,45,-26,-16,-16,-25,-12)

##source for paritisan lean: https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/
nc_congress_court$color <- case_when(
       nc_congress_court$Partisan_Lean  < -19 ~ -3,
        nc_congress_court$Partisan_Lean < -10 ~ -2,
        nc_congress_court$Partisan_Lean < -5 ~ -1,
        nc_congress_court$Partisan_Lean < 5 ~ 0,
        nc_congress_court$Partisan_Lean  < 10 ~ 1,
        nc_congress_court$Partisan_Lean > 10 ~ 2

)

nc_congress_court$Pred_Party <- case_when(
       nc_congress_court$Partisan_Lean  < 2 ~ 0,
        nc_congress_court$Partisan_Lean > 2 ~ 1
         
)




## You can use st_join and tag each census block centroid with the party ID from congressional layer,
## but it is more efficient to tag in a separate GIS software, then read back in already tagged. Work smarter not harder. 



```

Need to re-run this with the get_ACS

all blocks in 2020 census with:

Demographic: white, black, asian, hispanic
median income
Households



```{r message=FALSE, warning=FALSE, include=FALSE}
# census_api_key('ba3d95cae3658f3145daf7f66119ed58a52aaca3')
# 
# 
# #white = P1_003N
# #black = P1_004N
# #hispanic =P2_002N
# #asian = P1_006N
# #median income = "B19301_001
# 
# ## load data
# blocks_2020_White <- get_decennial(
#  geography="block",
#  variables= "P1_003N", 
#  summary_var = "P1_001N",
#  state="NC",
#  year=2020,
#  geometry=TRUE,
# )
# 
# 
# blocks_2020_White$PctWhite <- blocks_2020_White$value/blocks_2020_White$summary_value
# 
# 
# 
# blocks_2020_Black <- get_decennial(
#  geography="block",
#  variables= "P1_004N",
#  summary_var = "P1_001N",
#  state="NC",
#  #county="Durham",
#  year=2020,
#  geometry=TRUE
# )
# 
# blocks_2020_Black$PctBlack <- blocks_2020_Black$value/blocks_2020_Black$summary_value
# 
# 
# blocks_2020_Asian <- get_decennial(
#  geography="block",
#  variables= "P1_006N",
#  summary_var = "P1_001N",
#  state="NC",
#  year=2020,
#  geometry=TRUE
# )
# 
# blocks_2020_Asian$PctAsian <- blocks_2020_Asian$value/blocks_2020_Asian$summary_value
# 
# blocks_2020_Hispanic <- get_decennial(
#  geography="block",
#  variables= "P2_002N",
#  summary_var = "P2_001N",
#  state="NC",
#  year=2020,
#  geometry=TRUE
# )
# 
# blocks_2020_Hispanic$PctHispanic <- blocks_2020_Hispanic$value/blocks_2020_Hispanic$summary_value
# 
# 
# NC_Income <- get_acs(
#  geography="county",
#  variables= "B19301_001", #Median Income per Capita in 2019 Inflation Adjusted Dollars
#  state="NC",
#  year=2019,
#  geometry=TRUE,
# )  
# 
# 
# # 
# # final_bgs <- cbind(NC_Hisp_2020_bg, NC_black_2020_bg, NC_white_2020_bg, NC_HH_2020_bg,NC_Asian_2020_bg ) %>%
# #  select(GeoID, Name, pct_black, pct_hispanic, pct_white, pct_asian, HH_Vacant, geometry) %>%
# #  mutate(pct_other = 100 - (pct_black + pct_white + pct_asian))
# # 
# 
# blocks_2020 <- cbind(blocks_2020_White,blocks_2020_Black,blocks_2020_Asian,blocks_2020_Hispanic) %>%
#   select(GEOID, NAME, PctWhite, PctBlack, PctAsian, PctHispanic) 
# 
# temp <- st_transform(blocks_2020, st_crs(nc_congressional_2019)) %>% st_make_valid()
# 
# 
# centroids <- blocks_2020 %>%
#   st_point_on_surface() %>%
#   st_join(NC_Income %>%select(estimate)) %>%
#   mutate(Med_Income = estimate)
# 
# ##next step includes tagging each of the four political representation 
# 
# #first, tag existing 2019 representation
# 
# centroids <- st_transform(centroids, st_crs(trm_taz_g2)) %>% st_make_valid()
# nc_congressional_2019 <- st_transform(nc_congressional_2019, st_crs(trm_taz_g2)) %>% st_make_valid()
# nc_congressional_2021 <- st_transform(nc_congressional_2021, st_crs(trm_taz_g2)) %>% st_make_valid()
# nc_congressional_2021_feb <- st_transform(nc_congressional_2021_feb, st_crs(trm_taz_g2)) %>% st_make_valid()
# nc_congress_court <- st_transform(nc_congress_court, st_crs(trm_taz_g2)) %>% st_make_valid()
# 
# nc_congressional_2019 <- nc_congressional_2019 %>%
#   mutate( Party_2019 = case_when(
#     sParty == "Democrat" ~ 1,
#     sParty == "Republican" ~ 0
#   ))
# 
# 
# 
# centroids <- centroids %>%
#   st_join(nc_congressional_2019 %>% select(Party_2019 = sParty)) %>%
#   st_join(nc_congressional_2021 %>% select(Party_2021 = Pred_Party)) %>%
#   st_join(nc_congressional_2021_feb %>% select(Party_2021_Feb = Pred_Party)) %>%
#   st_join(nc_congress_court %>% select(Party_2021_courts = Pred_Party))
# 
# 
# st_write(centroids, "centroids.shp")

centroids <- st_read("/Users/jacobford/Desktop/centroids.shp")

 
```


## Visualize new maps

The state house and state senate maps are final, only the congressional maps are being altered. I didn't know that until I added the new maps (February in click option) to the below. If you don't believe me, just find out for yourself. 

Recall, to train the models on the new congressional lines, I assumed the 2022 midterms would follow FiveThirtyEight's [Partisan Lean](https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/). This assumes the worst for Democrats in North Carolina, which I imagine is more likely than not. 

```{r echo=FALSE, message=FALSE, warning=FALSE}




 congress_2019 <- st_transform(nc_congressional_2019, st_crs(trm_taz_g2)) %>% st_make_valid()

pal <- colorFactor(
  palette=c('blue','red'),
  domain=congress_2019$sParty
)



congress_2021 <- st_transform(nc_congressional_2021, st_crs(trm_taz_g2)) %>% st_make_valid()

congress_2021_Feb <- st_transform(nc_congressional_2021_feb, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congress_court <- st_transform(nc_congress_court, st_crs(trm_taz_g2)) %>% st_make_valid()

library(RColorBrewer)

colfunc <- colorRampPalette(c("#2C77BF", "red"))




pal_2021 <- colorFactor(
  palette=c('darkred','red', 'pink', 'grey', 'blue', 'darkblue'),
  domain=c(-3,-2,-1,0,1,2)
)



map <- leaflet(congress_2019) %>%
  addProviderTiles("CartoDB.Positron") %>%
  ##Congress
  addPolygons(data=congress_2019, group="Congress Old",
              popup=paste("Congressional District: ", congress_2019$District, "<br>",
                          "Party: ", congress_2019$sParty, "<br>",
                          "Representative 2019: ", congress_2019$sName),
              weight=1,
              color =~pal(congress_2019$sParty)) %>%
  addPolygons(data=congress_2021, group="Congress New",
              popup=paste("Congressional District: ", congress_2021$DISTRICT, "<br>", "Partisan Lean: ", congress_2021$Partisan_Lean ),dashArray="3", color =~pal_2021(congress_2021$color),
              weight=1, fillOpacity =0.5) %>%
  
  addPolygons(data=congress_2021_Feb, group="Congress February",
              popup=paste("Congressional District: ", congress_2021_Feb$DISTRICT, "<br>", "Partisan Lean: ", congress_2021_Feb$Partisan_Lean ),dashArray="3", color =~pal_2021(congress_2021_Feb$color),
              weight=1, fillOpacity =0.5) %>%

  addPolygons(data=nc_congress_court, group="Court Ordered Lines",
              popup=paste("Congressional District: ", nc_congress_court$DISTRICT, "<br>", "Partisan Lean: ", nc_congress_court$Partisan_Lean ),dashArray="3", color =~pal_2021(nc_congress_court$color),
              weight=1, fillOpacity =0.5) %>%
  
  
  
  addLayersControl(overlayGroups = c("Congress Old", "Congress New", "Congress February", "Court Ordered Lines"))

map %>% hideGroup("Congress New") %>% hideGroup("House February")%>% hideGroup("Congress February")



  
 


```















## Summary Stats

We'll make the same summary stats to start. Notice how the February updated lines seem to be a large improvement in racial parity - Blacks were nearly twice as likely to be represented by a Democrat with the 2010-2019 congressional lines, in the February update it's almost even. This process requires tagging each of the nearly quarter million census blocks in North Carolina to the congressional district represented.


```{r echo=FALSE, message=FALSE, warning=FALSE}


##rename

centroids$Party_2019 <- centroids$Pr_2019
centroids$Party_2021 <- centroids$Pr_2021
centroids$Party_2021_Feb <- centroids$P_2021_F
centroids$Party_2021_courts <- centroids$Pr_2021_ 

centroids$Med_Income <- centroids$Md_Incm
centroids$PctWhite <- centroids$PctWhit
centroids$PctBlack <- centroids$PctBlck
centroids$PctHispanic <- centroids$PctHspn
centroids$PctAsian <- centroids$PctAsin
  ###
  
  
centroids <- centroids %>%
  filter(!is.na(Party_2019))

centroids$Med_income_log <- case_when(
  centroids$Med_Income == 0 ~ 0,
  centroids$Med_Income > 0 ~ log(centroids$Med_Income))

## 2019 
temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`PARTY 2019` =Party_2019) %>%
  mutate('Party_2019_tag' = case_when(
    Party_2019=="Democrat" ~ 1,
    TRUE ~ 0
    
  )) %>%
  group_by(`PARTY 2019`) %>%
  summarize(Mean_white = mean(PctWhite, na.rm=T),
            Mean_black = mean(PctBlack, na.rm=T),
            Mean_hispanic = mean(PctHispanic, na.rm=T),
            Mean_asian = mean(PctAsian, na.rm=T),
            Mean_inc = mean(Med_Income, na.rm=T))

library(DT)

datatable(temp, caption = "Average Racial Breakdown by 2019 Congressional Representation") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)

## 2020

temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`PARTY 2021` = case_when(
    Party_2021 == 0 ~ "Republican",
    Party_2021 == 1 ~ "Democrat"
  )) %>%
  group_by(`PARTY 2021`) %>%
  summarize(Mean_white = mean(PctWhite, na.rm=T),
            Mean_black = mean(PctBlack, na.rm=T),
            Mean_hispanic = mean(PctHispanic, na.rm=T),
            Mean_asian = mean(PctAsian, na.rm=T),
            Mean_inc = mean(Med_Income, na.rm=T))

library(DT)

datatable(temp, caption = "Average Racial Breakdown by 2021 Congressional Representation") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)



## February

temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`February_Party` = case_when(
    Party_2021_Feb == 0 ~ "Republican",
    Party_2021_Feb == 1 ~ "Democrat"
  )) %>%
  group_by(`February_Party`) %>%
  summarize(Mean_white = mean(PctWhite, na.rm=T),
            Mean_black = mean(PctBlack, na.rm=T),
            Mean_hispanic = mean(PctHispanic, na.rm=T),
            Mean_asian = mean(PctAsian, na.rm=T),
            Mean_inc = mean(Med_Income, na.rm=T))
library(DT)

datatable(temp, caption = "February Update") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)


## court Ordered

temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`Court Predicted Part` = case_when(
    Party_2021_courts == 0 ~ "Republican",
    Party_2021_courts == 1 ~ "Democrat"
  )) %>%
  group_by(`Court Predicted Part`) %>%
  summarize(Mean_white = mean(PctWhite, na.rm=T),
            Mean_black = mean(PctBlack, na.rm=T),
            Mean_hispanic = mean(PctHispanic, na.rm=T),
            Mean_asian = mean(PctAsian, na.rm=T),
            Mean_inc = mean(Med_Income, na.rm=T))

library(DT)

datatable(temp, caption = "Court Ordered Maps") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)

```












```{r message=FALSE, warning=FALSE, include=FALSE}

##Train and Test Data Sets
#replace na's w/ zeros

centroids <- centroids %>%
  mutate( Party_2019_dummy = case_when(
    Party_2019 == "Democrat" ~ 1,
    Party_2019 == "Republican" ~ 0
  ))

centroids[is.na(centroids)]=0
centroids <- centroids %>% st_drop_geometry()
train <- centroids[1:177478,]
test <- centroids[177479:236638,]

```




```{r message=FALSE, warning=FALSE, include=FALSE}

### 2019 Model Outputs



model_2019_blocks <- glm(Party_2019_dummy ~PctHispanic + PctWhite + PctBlack + PctAsian + Med_income_log ,family=binomial(link='logit'),data=train)

summary(model_2019_blocks)





```


```{r message=FALSE, warning=FALSE, include=FALSE}
### 2021 Model Outputs


model_2021 <- glm(Party_2021 ~PctHispanic + PctWhite + PctBlack + PctAsian + Med_income_log ,family=binomial(link='logit'),data=train)

summary(model_2021)



```

```{r message=FALSE, warning=FALSE, include=FALSE}
### February Model


feb_model <- glm(Party_2021_Feb ~PctHispanic + PctWhite + PctBlack + PctAsian + Med_income_log ,family=binomial(link='logit'),data=train)

summary(feb_model)



```


```{r message=FALSE, warning=FALSE, include=FALSE}
### Court Ordered Model


court_model <- glm(Party_2021_courts ~PctHispanic + PctWhite + PctBlack + PctAsian + Med_income_log ,family=binomial(link='logit'),data=train)


summary(court_model)



```




## Logistic Regression Curve

Lets see if the new lines hold up with the logistic model, determining how likely a particular demographic is to be represented by a Democrat (note 1 - the y-axis = probability Republican). Black and Asian populations are more likely to be represented by Republicans in large numbers in the February update. Note how at the very end of the logistic curve, in those census blocks with 90, 95% racial homogeneity, the likelihood of being represented by a democrat is far less likely compared to the 2010-2019 or 2021 proposed districts. This only holds for black populations as you can see in the below: 

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(test, aes(x=PctAsian, y=Party_2019_dummy)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Asian")

ggplot(test, aes(x=PctAsian, y=Party_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Asian")

ggplot(test, aes(x=PctAsian, y=Party_2021_Feb)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

ggplot(test, aes(x=PctAsian, y=Party_2021_courts)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("Court Ordered Update")

#=====================================================================#


ggplot(test, aes(x=PctBlack, y=Party_2019_dummy)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Black")

ggplot(test, aes(x=PctBlack, y=Party_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Black")


ggplot(test, aes(x=PctBlack, y=Party_2021_Feb)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

ggplot(test, aes(x=PctBlack, y=Party_2021_courts)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("Court Ordered Update")

#=====================================================================#
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(test, aes(x=PctWhite, y=Party_2019_dummy)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts White")


ggplot(test, aes(x=PctWhite, y=Party_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts White")

ggplot(test, aes(x=PctWhite, y=Party_2021_Feb)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

ggplot(test, aes(x=PctWhite, y=Party_2021_courts)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("Court Ordered Update")
#=====================================================================#

ggplot(test, aes(x=PctHispanic, y=Party_2019_dummy)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Hispanic")

ggplot(test, aes(x=PctHispanic, y=Party_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Hispanic")

ggplot(test, aes(x=PctHispanic, y=Party_2021_Feb)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

ggplot(test, aes(x=PctHispanic, y=Party_2021_courts)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("Court Ordered Update")

#=====================================================================#




```


## Model Fits

The logistic model fit in the 2010-19 districts was a paltry 71.1%, drastically improved upon by the 2021 struck-down lines - up to 88.3%. The February lines are in the middle, at 82.6%; meaning this simple logistic model, fed with nearly a quarter million census block values, developed into a machine learning algorithm that could predict with 82.6% accuracy the political representation of a district based solely on racial makeup. 



```{r echo=FALSE, message=FALSE, warning=FALSE}

## Predictive Fits

pred <- predict(model_2019_blocks, newdata = test, type = "response")



y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$Party_2019_dummy

pred_2019 <- mean(y_pred == y_act, na.rm=T) 

#=====================================================================#

pred <- predict(model_2021, newdata = test, type = "response")

y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$Party_2021

pred_2021 <- mean(y_pred == y_act, na.rm=T) 





#=====================================================================#
pred <- predict(feb_model, newdata = test, type = "response")

y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$Party_2021

feb_pred <- mean(y_pred == y_act, na.rm=T) 
#=====================================================================#



#=====================================================================#
pred <- predict(court_model, newdata = test, type = "response")

y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$Party_2021_courts

court_pred <- mean(y_pred == y_act, na.rm=T) 
#=====================================================================#



vals <- c(pred_2019, pred_2021,feb_pred,court_pred)





df <- data.frame(Year = c('2019', '2021', 'February', 'Court Ordered'),vals)

df$vals <- round(df$vals, 3)



ggplot(df, aes(x=Year, y=vals))+geom_bar(stat='identity', position='dodge') + geom_text(aes(label=vals), vjust=-0.25)+
  xlab("Year")+ylab("Predictive Strength")+theme_minimal()+ggtitle("Logit Model Accuracy")


```





```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

## Odds Ratios
exp(coef(model_2019_blocks))
exp(coef(model_2021))
exp(coef(feb_model))
exp(coef(court_model))

exp(cbind(OR = coef(model_2019_blocks), confint(model_2019_blocks)))
exp(cbind(OR = coef(model_2021), confint(model_2021)))
exp(cbind(OR = coef(feb_model), confint(feb_model)))
exp(cbind(OR = coef(court_model), confint(court_model)))


```




## Odds Ratio Box Plot


```{r echo=FALSE, message=FALSE, warning=FALSE}
boxLabels = c("Hispanic", "White", "Black", "Asian")

##2019


df <- data.frame(lables=boxLabels, 
                  boxOdds = c(1.0021304779, 0.9934778163  , 1.0200081806 ,1.0312674226    ), 
                        boxCILow = c(1.0014661525   , 0.9931773600   ,1.0195372923 ,1.0295998976     ), 
                        boxCIHigh = c(1.002793072, 0.993778233, 1.020480101,1.032950946))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("2019 Model Odds Ratios, Race")


##2021


df <- data.frame(lables=boxLabels, boxOdds = c(1.012368848609690   , 
0.993498405513978   , 1.013385901077802,1.042262592003207    ), boxCILow = c(1.011481634734796   , 0.993044840129250   ,1.012728356907708,1.040464698203379   ), 
                        boxCIHigh = c(1.013251325461511, 0.993952139817316, 1.014042344750129,1.044072940422991))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("2021 Model Odds Ratios, Race")



##February


df <- data.frame(lables=boxLabels, boxOdds = c(1.00953308    , 
0.99998742    , 1.00236907 ,1.02191628     ), boxCILow = c(1.0088145    , 0.9996327    ,1.0018087 ,1.0203451    ), 
                        boxCIHigh = c(1.01024899, 1.00034257, 1.00292776,1.02348719))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("February Model Odds Ratios, Race")


##Court Ordered


df <- data.frame(lables=boxLabels, boxOdds = c(1.0102481962     , 
0.9986372917     , 1.0057546861  ,1.0135513824      ), boxCILow = c(1.0094911593     , 0.9982595907     ,1.0051712000  ,1.0118822905     ), 
                        boxCIHigh = c(1.011002040, 0.999015431, 1.006336640,1.015208764))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("Court Ordered Model Odds Ratios, Race")









# Median Income


boxLabels = c("2019 District", "2021 Districts", "February", "Court Ordered")

df <- data.frame(lables=boxLabels, boxOdds = c(2.0259539656   , 5.310856286752776 , 1.18227618 ,1.6628874988    ), boxCILow = c(1.9723423285   , 5.108356308166039 ,1.1563041,1.6096197402    ), 
                        boxCIHigh = c(2.081119204,5.521796860873518,1.20975856,1.718109910))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("Median Income Logged")




```


## Random Forest 


```{r echo=FALSE, message=FALSE, warning=FALSE}

model_data <- centroids %>% 
  select(Party_2019_dummy,Party_2021,Party_2021_Feb, Party_2021_courts, PCT_BLC=PctBlack, PCT_HSP=PctHispanic, PCT_WHT=PctWhite, PCT_ASN=PctAsian, Med_income_log)
model_data$PARTY_2021 <- as.factor(model_data$Party_2021)
model_data$PARTY_2019 <- as.factor(model_data$Party_2019_dummy)
model_data$PARTY_2022 <- as.factor(model_data$Party_2021_Feb)
model_data$COURT_PART <- as.factor(model_data$Party_2021_courts)

samp <- sample(nrow(model_data), 0.8*nrow(model_data))

train <- model_data[samp,]
test <- model_data[-samp,]

RF_2019 <- randomForest(
  formula = PARTY_2019 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2019)
varImpPlot(RF_2019)
#=====================================================================#


RF_2021<- randomForest(
  formula = PARTY_2021 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2021)
varImpPlot(RF_2021)
#=====================================================================#


RF_Feb<- randomForest(
  formula = PARTY_2022 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2021)
varImpPlot(RF_Feb)


#=====================================================================#


court_Feb<- randomForest(
  formula = COURT_PART ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2021)
varImpPlot(court_Feb)










```




Finally, lets see the predictive fit for all four: 



```{r echo=FALSE, message=FALSE, warning=FALSE}



p <- predict(RF_2019, newdata=test)
# table(p, test$PARTY_2019)

results <- cbind(p, test$PARTY_2019)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_2019 <- sum(p==test$PARTY_2019) / nrow(test)


#=====================================================================#



p <- predict(RF_2021, newdata=test)
# table(p, test$PARTY_2021)

results <- cbind(p, test$PARTY_2021)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_2021 <- sum(p==test$PARTY_2021) / nrow(test)


#=====================================================================#


p <- predict(RF_Feb, newdata=test)
# table(p, test$PARTY_2021)

results <- cbind(p, test$PARTY_2022)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_feb <- sum(p==test$PARTY_2022) / nrow(test)


#=====================================================================#
p <- predict(court_Feb, newdata=test)
# table(p, test$PARTY_2021)

results <- cbind(p, test$COURT_PART)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_court <- sum(p==test$COURT_PART) / nrow(test)


#=====================================================================#



vals <- c(pred_2019, pred_2021,pred_feb,pred_court)

df <- data.frame(Year = c('2019', '2021', 'February', 'Court Ordered'),vals)

df$vals <- round(df$vals, 3)



ggplot(df, aes(x=Year, y=vals))+geom_bar(stat='identity', position='dodge') + geom_text(aes(label=vals), vjust=-0.25)+
  xlab("Year")+ylab("Predictive Strength")+theme_minimal()+ggtitle("Random Forest Model Accuracy")

```




## Conclusion








